---
format: 
  html:
    self-contained: true
    css: 
      - ../assets/webex.css
    include-after-body: 
      - ../assets/webex.js
---

```{r}
#| include: false
library(webexercises)
```

**Download the data**



Filename |  Description | Download | Info
------ | -------------|  ---: | ---:
`roadside-transects.csv` | Lowland-to-highland Roadside Transects Data, South-East Highlands, Australia, 2017. | <a href='data/roadside-transects.csv' download>{{< fa file-csv >}}</a> | [{{< fa info-circle >}}](https://datacommons.anu.edu.au/DataCommons/rest/display/anudc:5861?layout=def:display){target="_blank"}
`branch-water-derived-value.xlsx` | The effect of salinity on branch water relations and stem hydraulic vulnerability in two co-occurring mangrove species. | <a href='data/branch-water-derived-value.xlsx' download>{{< fa file-excel >}}</a> | [{{< fa info-circle >}}](https://datacommons.anu.edu.au/DataCommons/rest/display/anudc:6254?layout=def%3adisplay){target="_blank"}




Download the above datasets.

```{r}
#| include: false
branch <- readxl::read_excel("data/branch-water-derived-value.xlsx", sheet = "Values for RWC figures")
branchkey <- readxl::read_excel("data/branch-water-derived-value.xlsx", sheet = "Key Values")
roadside <- readr::read_csv("data/roadside-transects.csv")
```


- Read the data in `roadside-transects.csv`. How many variables does this sheet consist of? `r fitb(ncol(roadside))` How many observations are there? `r fitb(nrow(roadside))`
- Read the sheet "Values for RWC figures" in `branch-water-derived-value.xlsx`. How many variables does this sheet consist of? `r fitb(ncol(branch))` How many observations are there? `r fitb(nrow(branch))`
- Read the sheet "Key Values" in `branch-water-derived-value.xlsx`. How many variables does this sheet consist of? `r fitb(ncol(branchkey))` How many observations are there? `r fitb(nrow(branchkey))`